Linear Algebra as a Unifying Theory\\of Wireless Workload Modeling

### 2008-02-02 paper.tex
Just realized that protocol type is not represented. Crap.

The traffic workload pattern in a network is simply a collection of packets between hosts in the network sent at certain times. It is standard in network analysis, however, to aggregate sequences of packets sharing the same \caps{IP} protocol, source and destination nodes, and source and destination \caps{TCP} or \caps{UDP} port numbers. Such a sequence of packets is called a \textit{flow}. The traffic pattern is characterized completely by the detailed behavior of it's flows. The behavior of a flow is defined by seven properties: its source and destination nodes; its start and end times; the specific sizes of and intervals between its constituent packets; and finally, whether it is a response, paired with another flow.\footnote{Flow pairing is typically not addressed in traffic models. However, the most common transport protocol, {\scriptsize{TCP}}, is full-duplex, exhibiting tightly coupled pairs of client-server flows. Therefore, accurate representation of traffic behavior requires knowing which flows are request-response pairs. Moreover, although they are not strictly full-duplex, it is generally appropriate to view {\scriptsize{UDP}} and even {\scriptsize{ICMP}} flows as paired, since they are typically paired. For example, {\scriptsize{DNS}} responses are naturally associated with the corresponding request, and will have symmetrical {\scriptsize{IP}} addresses and port numbers.} We propose to represent the behavior of each flow in terms of seven vectors---one for each aspect of its behavior. The aggregate behavior of all the flows in a workload pattern may then be represented by these matrices, the rows of which are the behavior vectors of the flows.

@@

The {packets-bytes-duration} model representation is clearly overly simplistic and very incomplete. (Yet, it is more detailed than most commonly used experimental traffic models!) It captures nothing about variability of packet sizes, nor about the ``burstiness'' of packet transmissions; it leaves start time unspecified; it says nothing about the source or destination of the flow. At the opposite extreme, it is possible to embed the exact start time of each flow, and the exact size and transmission times of its packets in a vector space, etc. The difficulty comes in doing so in a manner that makes the summation of vectors correspond naturally and meaningfully to the aggregate behavior of flows. We take an intermediate approach, suggested by the results of~\cite{Karpinski07:realism}: to reproduce network performance accurately, it suffices to know, for each flow, the correct source node, destination node, start time, end time, empirical distributions of packet sizes and inter-packet intervals for each flow, and which other flow (if any) it is paired with. If there are $f$ flows in the entire workload pattern, then its collective behavior can be described with seven matrices; the rows of each matrix are behavior vectors of all flows, indexed in some chosen order. We use the following notation for the behavior matrices and their constituent vectors:\vspace{-0.25em}
\begin{tabbing}
\hspace{1.5em}\=\bullet\hspace{0.5em}\=\hspace{9.75em}\=
$D\:$\=$=\setx{\vec{d}_i}\:$\=$=\setx{d_{ij}}\:$\=\kill
\>\bullet\> sources: \>$S$\>$=\setx{\vec{s}_i}$\>$=\setx{s_{ij}}$\>$\in\R^{f \times n}$,\\
\>\bullet\> destinations: \>$D$\>$=\setx{\vec{d}_i}$\>$=\setx{d_{ij}}$\>$\in\R^{f \times n}$,\\
\>\bullet\> flow pairs: \>$P$\>$=\setx{\vec{p}_i}$\>$=\setx{p_{ij}}$\>$\in\R^{f \times f}$,\\
\>\bullet\> start times: \>$B$\>$=\setx{\vec{b}_i}$\>$=\setx{b_{ij}}$\>$\in\R^{f \times d_t}$,\\
\>\bullet\> end times: \>$E$\>$=\setx{\vec{e}_i}$\>$=\setx{e_{ij}}$\>$\in\R^{f \times d_t}$,\\
\>\bullet\> packet sizes: \>$Z$\>$=\setx{\vec{z}_i}$\>$=\setx{z_{ij}}$\>$\in\R^{f \times d_z}$,\\
\>\bullet\> inter-packet intervals: \>$V$\>$=\setx{\vec{v}_i}$\>$=\setx{v_{ij}}$\>$\in\R^{f \times d_v}$.
\end{tabbing}

@@

\subsection{Flow Pairing: Clients \& Servers}

>>>>

The traffic workload pattern in a network is simply a collection of packets between hosts in the network sent at certain times. It is standard in network analysis, however, to aggregate sequences of packets sharing the same \caps{IP} protocol, source and destination nodes, and source and destination \caps{TCP} or \caps{UDP} port numbers. Such a sequence of packets is called a \textit{flow}. The traffic pattern is characterized completely by the detailed behavior of it's flows. The behavior of a flow is defined by seven properties: its source and destination nodes; its start and end times; the specific sizes of and intervals between its constituent packets; and finally, whether it is paired with another flow in a request-response relationship.\footnote{Flow pairing is typically not addressed in traffic models. However, the most common transport protocol, {\scriptsize{TCP}}, is full-duplex, exhibiting tightly coupled pairs of client-server flows. Therefore, accurate representation of traffic behavior requires knowing which flows are request-response pairs. Moreover, although they are not strictly full-duplex, it is generally appropriate to view {\scriptsize{UDP}} and {\scriptsize{ICMP}} flows as paired, since the client-server paradigm is still used. Each {\scriptsize{DNS}} response, for example, is naturally associated with its corresponding request, and will have symmetrical {\scriptsize{IP}} addresses and port numbers.} We propose to represent the behavior of each flow in terms of seven vectors---one for each aspect of its behavior. The aggregate behavior of all the flows in a workload pattern are then represented as matrices, the rows of which are the behavior vectors of the flows.

@@

The {packets-bytes-duration} representation is clearly overly simplistic and very incomplete. (Yet, it is more detailed than the most commonly used experimental traffic models!) It captures nothing about variability of packet sizes, nor about the ``burstiness'' of packet transmissions; it leaves start time unspecified; it says nothing about the source or destination of the flow. At the opposite extreme, it is possible to embed the exact start time of each flow, and the exact size and transmission times of its packets in a vector space, etc. The difficulty comes in doing so in a manner that makes the summation of vectors correspond naturally and meaningfully to the aggregate behavior of flows. We take an intermediate approach, suggested by the results of~\cite{Karpinski07:realism}: to reproduce network performance accurately, it suffices to know, for each flow, the correct source node, destination node, start time, end time, empirical distributions of packet sizes and inter-packet intervals for each flow, and which other flow (if any) it is a response to. If there are $f$ flows in the entire workload pattern, then its collective behavior can be described with seven matrices; the rows of each matrix are the behavior vectors of the flows, indexed in some chosen order. We use the following notation for the behavior matrices and their constituent vectors and elements:\vspace{-0.25em}
\begin{tabbing}
\hspace{1.5em}\=\bullet\hspace{0.5em}\=\hspace{9.75em}\=
$D\:$\=$=\setx{\vec{d}_i}\:$\=$=\setx{d_{ij}}\:$\=\kill
\>\bullet\> sources: \>$S$\>$=\setx{\vec{s}_i}$\>$=\setx{s_{ij}}$\>$\in\R^{f \times n}$,\\
\>\bullet\> destinations: \>$D$\>$=\setx{\vec{d}_i}$\>$=\setx{d_{ij}}$\>$\in\R^{f \times n}$,\\
\>\bullet\> flow pairs: \>$P$\>$=\setx{\vec{p}_i}$\>$=\setx{p_{ij}}$\>$\in\R^{f \times f}$,\\
\>\bullet\> start times: \>$B$\>$=\setx{\vec{b}_i}$\>$=\setx{b_{ij}}$\>$\in\R^{f \times d_t}$,\\
\>\bullet\> end times: \>$E$\>$=\setx{\vec{e}_i}$\>$=\setx{e_{ij}}$\>$\in\R^{f \times d_t}$,\\
\>\bullet\> packet sizes: \>$Z$\>$=\setx{\vec{z}_i}$\>$=\setx{z_{ij}}$\>$\in\R^{f \times d_z}$,\\
\>\bullet\> inter-packet intervals: \>$V$\>$=\setx{\vec{v}_i}$\>$=\setx{v_{ij}}$\>$\in\R^{f \times d_v}$.
\end{tabbing}

@@

\subsection{Flow Pairing: Requests \& Responses}
